Throughput Prediction on 60 GHz Devices for Latency Sensitive Applications

### 2021-01-21 paper.tex
changes

Due to the recent advancements in multimedia technology, video content is getting richer with higher resolution and better quality. However, this also means that applications that provide such content demand much more bandwidth from the underlying network. For example, most high quality VR setups today are tethered because today's wireless networks are unable to provide the necessary capacity to support such applications. However, with the advent of mmWave technologies like 5G (for cellular networks) and 802.11ad (for indoor WLANs) these applications can indeed go wireless. For example, the IEEE 802.11ad wLAN standard ~\cite{80211ad} governs the use of the 14 GHz of unlicensed spectrum around 60 GHz. The standard supports 2 GHz wide channel and provides PHY data rates of up to 6.7 Gbps which is ideal to satisfy the needs for such applications.

However, due to their much higher frequency of operation, 60 GHz networks have vastly different propagation characteristics from sub-6 GHz networks. Due to the high attenuation loss at 60 GHz, directional communication is needed. This makes the wireless link highly susceptible to human blockage and mobility. Due to these challenges, a user watching a 360\degree~video or playing a VR game over a 60 GHz network can experience large periods of rebuffering/stalls due to intervals of low to no 60 GHz connectivity. For example, the user may be moving around or rotating in such a way so as to face completely away from the access point and thus self-block the link. Hence, even if 60 GHz wireless is used to enable these high resolution applications, it cannot be used as a standalone technology. One way to keep the rebuffering and stalls to a minimum while using 60 GHz is to use legacy sub-6 GHz WiFi, which does not suffer from these issues of blockage and mobility, as a backup. %This means that there will be a drop in content quality due to the use of slower sub-6 GHz network but will still lead to better user QoE than when there are stalls. 
These unique characteristics of the 60 GHz links make throughput prediction in 60 GHz WLANs a much more important and challenging problem that in legacy WLANs.

In this work, \textit{for first time}, we study the throughput predictability in 60 GHz wireless LANs using machine learning. In the past, throughput estimation/prediction has been studied in the context of sub-6 GHz wireless networks~\cite{khan:infocom2016,song:secon2017,kajita:wmnc2015} and mobile networks~\cite{liu:globecom2015}. However there has been no previous work, to our best knowledge, which has studied throughput prediction at mmWave frequencies such as 60 GHz.

>>>>

Due to the recent advancements in multimedia technology, video content
is getting richer with higher resolution and better quality. However,
this also means that applications that provide such content demand
much more bandwidth from the underlying network. For example, most
high quality VR setups today are tethered because today's wireless
networks are unable to provide the necessary capacity to support such
applications. However, with the advent of mmWave technologies like 5G
(for cellular networks) and 802.11ad (for indoor WLANs), these
applications can potentially go wireless. For example, the IEEE 802.11ad
wLAN standard~\cite{80211ad} governs the use of the 14 GHz of
unlicensed spectrum around 60 GHz. The standard supports 2 GHz wide
channel and provides PHY data rates of up to 6.7 Gbps which is ideal
to satisfy the needs for such applications.

However, due to their much higher frequency of operation, 60 GHz networks have vastly different propagation characteristics from sub-6 GHz networks. Due to the high attenuation loss at 60 GHz, directional communication is needed. This makes the wireless link highly susceptible to human blockage and mobility. Due to these challenges, a user watching a 360\degree~video or playing a VR game over a 60 GHz network can experience large periods of rebuffering/stalls due to intervals of low to no 60 GHz connectivity. For example, the user may be moving around or rotating in such a way so as to face completely away from the access point and thus self-block the link. Hence, even if 60 GHz wireless is used to enable these high resolution applications, it cannot be used as a standalone technology. One way to keep the rebuffering and stalls to a minimum while using 60 GHz is to use legacy sub-6 GHz WiFi, which does not suffer from these issues of blockage and mobility, as a backup.
 %This means that there will be a drop in content quality due to the use of slower sub-6 GHz ne% twork but will still lead to better user QoE than when there are stalls. 
These unique characteristics of the 60 GHz links make throughput prediction in 60 GHz WLANs a much more important and challenging problem than in legacy WLANs.

In this work, {for the first time}, we study the throughput predictability in 60 GHz wireless LANs using machine learning. In the past, throughput estimation/prediction has been studied in the context of sub-6 GHz wireless networks~\cite{khan:infocom2016,song:secon2017,kajita:wmnc2015} and mobile networks~\cite{liu:globecom2015}. However there has been no previous work, to our best knowledge, which has studied throughput prediction at mmWave frequencies such as 60 GHz.
